---
title: "Customer Retention"
author: "Lirong Wei"
date: "February 24, 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data
Telco Customer Churn (Focused customer retention programs)
https://www.kaggle.com/blastchar/telco-customer-churn

This is a small sample dataset (172KB) about telecom customers. There are 7043 observations and 21 variables and only 11 missing values. Each observation is a customer and the 21 variables include information about customer churn, services customer signed, customer account and demographic information. 
## Question of interest

The most interest questions in this datasets would be how to predict whether a customer churn or not and what are most important factors/predictors/variables that are related with customer churn? If we could provide churn customers before they actually leave, then we could privide some deal or attractive offer to retain our customers. 


## Approach to solve the problem

To predict customer churn, the variable churn would be our response variable. Churn only has two possible values: Yes or No. We could model a customer churning as a  Bernoulli random variable. Take all other information we have as covariates to explore what covariate will influent customer churn and to do prediction. Therefore, I choose logistics regression model:
$Churn_{i} ~ Bernoulli(\pi_{i})$

$Churn = Yes, with probability \pi_{i} $

$logit(\pi_{i})=\beta_{0} + \beta_{1}X_{i1} + \beta_{2}X_{i2} + ... +  \beta_{p-1}X_{i(p-1_)} $


## Code and results

Here is the dataset:
```{r loaddata}
cus_retention <- read.csv("WA_Fn-UseC_-Telco-Customer-Churn.csv")
 str(cus_retention)
head(cus_retention)
#check missing values
sum(is.na(cus_retention))
```

summary of the data:
```{r data_explore}
summary(cus_retention[,-1])
# We could easily find that PhoneService and MultipleLines must be perfectly collinear
# In addition, variable InternetService must cause collinearity among variables:  OnlineSecurity, OnlineBackup, DeviceProtection, TechSupport,  StreamingTV and StreamingMovies. 

#We could deal with this using data processing
cus_retention_new <-  cus_retention[,-1] #delete ID column
#1.Dummy variable for MultipleLines
cus_retention_new$MultipleLines_new <- 1
cus_retention_new$MultipleLines_new[cus_retention_new$MultipleLines=="No" |cus_retention_new$MultipleLines=="No phone service" ] <- 0
#2.Dummy variable for OnlineSecurity
cus_retention_new$OnlineSecurity_new <- 1
cus_retention_new$OnlineSecurity_new[cus_retention_new$OnlineSecurity=="No" |cus_retention_new$OnlineSecurity=="No internet service" ] <- 0
#3.Dummy variable for OnlineBackup
cus_retention_new$OnlineBackup_new <- 1
cus_retention_new$OnlineBackup_new[cus_retention_new$OnlineBackup=="No" |cus_retention_new$OnlineBackup=="No internet service" ] <- 0
#4.Dummy variable for DeviceProtection
cus_retention_new$DeviceProtection_new <- 1
cus_retention_new$DeviceProtection_new[cus_retention_new$DeviceProtection=="No" |cus_retention_new$DeviceProtection=="No internet service" ] <- 0
#5.Dummy variable for TechSupport
cus_retention_new$TechSupport_new <- 1
cus_retention_new$TechSupport_new[cus_retention_new$TechSupport=="No" |cus_retention_new$TechSupport=="No internet service" ] <- 0
#6.Dummy varible for StreamingTV
cus_retention_new$StreamingTV_new <- 1
cus_retention_new$StreamingTV_new[cus_retention_new$StreamingTV=="No" |cus_retention_new$StreamingTV=="No internet service" ] <- 0
#7.Dummy variable for StreamingMovies
cus_retention_new$StreamingMovies_new <- 1
cus_retention_new$StreamingMovies_new[cus_retention_new$StreamingMovies=="No" |cus_retention_new$StreamingMovies=="No internet service" ] <- 0

#Check process
table(cus_retention_new$MultipleLines_new)
table(cus_retention_new$OnlineSecurity_new)
table(cus_retention_new$OnlineBackup_new)
table(cus_retention_new$DeviceProtection_new)
table(cus_retention_new$TechSupport_new)
table(cus_retention_new$StreamingTV_new)
table(cus_retention_new$StreamingMovies_new)

#Ok! ready to build model
cus_retention_new <- cus_retention_new[,-c(7,9,10,11,12,13,14)]
```


Logistic regression model
```{r model,warning=FALSE}
library(caTools) #split data
library(ROCR) #roc curve and auc
library(MASS) #select variables 

#split dataset
set.seed(2019)
split = sample.split(cus_retention_new$Churn, SplitRatio = 0.7)
train = cus_retention_new[split,]
test = cus_retention_new[!split,]
#check split
table(train$Churn)/nrow(train)
table(test$Churn)/nrow(test)
#logistic regression model
logi_model <- glm(Churn ~ .,data = train,family = "binomial")
summary(logi_model)
```


```{r}
# Now we have 19 predictors we could use forward variable selection process to select variables
step <- stepAIC(logi_model,trace = FALSE)
summary(step)
#step is the final model.
#Here is the list of 12 varibles that are selected in the final model with point estimation of coeffient. They are significantly related with customer churn. 
step$coefficients
```


```{r evaluations}
# with the model, try to predict on test model
test$predProbs <- predict(step,newdata=test,type="response")
# evaluate with roc curve on test set
roc.pred = prediction(test$predProbs, test$Churn)
perf = performance(roc.pred, "tpr", "fpr")
plot(perf,                      # the data
     main = "ROC Curve",        # the chart's title
     xlab = "1 - Specificity",  # the name of the x-axis
     ylab = "Sensitivity",      # the name of the y-axis
     colorize=TRUE)             # add color to curve depending on cutoff prob.
# Select cutoff to be 0.3 as a tradeoff between sensitivity and specificity (similar to the original proportions of Churn in the dataset)
cutoff=0.3
test$predChurn = ifelse(test$predProbs >= cutoff, 1, 0)
#Confusion matrix
table(test$Churn, test$predChurn)
# Accuracy
(1174 + 438)/nrow(test)
# Sensitivity
438 / (123 + 438)
# Specificity
1174 / (1174 + 376)
#Lift curve
LiftData <- performance(roc.pred, measure="lift", x.measure="rpp")
plot(LiftData,                # the data
     main="Lift Chart",       # the chart's title
     xlab="% of Population",  # the name of the x-axis
     ylab="Lift",             # the name of the y-axis
     col="blue")              # the color of our lift curve

abline(1,0,col="red") # this adds a straight line at intercept 1 with slope 0
```

## Discoveries

1. Based on Z-test and variable selection process: Dependents, tenure, InternetService, Contract, PaperlessBilling, PaymentMethod, MonthlyCharges, TotalCharges, MultipleLines, OnlineSecurity, TechSupport, StreamingTV and StreamingMovies are significant predictors in predicting customer Churn. 

2. With Logistics regression we could predict probability of a customer Churn which will help us in customer detention. If we prioritize the correct prediction of Churn customers, we could try to use a threshold that will lead to high sensitivity of the model.

3. Could use machine learning method to do prediction and use cross-validation to compare the performance of models.

## Appendix: EDA figures

```{r echo=FALSE}
library(ggplot2)
library(dplyr)
plot0 <- cus_retention %>%
  group_by(Churn) %>%
  summarise(Customers=n())
ggplot(plot0) +
      geom_col(aes(x = Churn, y = Customers,fill=Churn))+
      labs(title = "Churn distribution")+
      theme_minimal()
  

plot1 <- cus_retention %>%
  group_by(gender,Churn) %>%
  summarise(Customers=n())
ggplot(plot1) +
      geom_col(aes(x = gender, y = Customers, fill = Churn))+
      labs(title = "Gender distribution")+
      theme_minimal()

plot2 <- cus_retention %>%
  group_by(Partner,Churn) %>%
  summarise(Customers=n())
ggplot(plot2) +
      geom_col(aes(x = Partner , y = Customers, fill = Churn))+
      labs(title = "Partner distribution")+
      theme_minimal()

plot3 <- cus_retention %>%
  group_by(Dependents,Churn) %>%
  summarise(Customers=n())
ggplot(plot3) +
      geom_col(aes(x = Dependents , y = Customers, fill = Churn))+
      labs(title = "Dependents distribution")+
      theme_minimal()

plot4 <- cus_retention %>%
  group_by(PhoneService,Churn) %>%
  summarise(Customers=n())
ggplot(plot4) +
      geom_col(aes(x = PhoneService , y = Customers, fill = Churn))+
      labs(title = "PhoneService distribution")+
      theme_minimal()

plot5 <- cus_retention %>%
  group_by(InternetService,Churn) %>%
  summarise(Customers=n())
ggplot(plot5) +
      geom_col(aes(x = InternetService , y = Customers, fill = Churn))+
      labs(title = "InternetService distribution")+
      theme_minimal()

plot6 <- cus_retention %>%
  group_by(PaperlessBilling,Churn) %>%
  summarise(Customers=n())
ggplot(plot6) +
      geom_col(aes(x = PaperlessBilling , y = Customers, fill = Churn))+
      labs(title = "PaperlessBilling distribution")+
      theme_minimal()

plot7 <- cus_retention %>%
  group_by(PaymentMethod,Churn) %>%
  summarise(Customers=n())
ggplot(plot7) +
      geom_col(aes(x = PaymentMethod , y = Customers, fill = Churn))+
      labs(title = "PaymentMethod distribution")+
      theme_minimal()

ggplot(cus_retention, aes(x=Churn, y=MonthlyCharges, fill=Churn)) +
  geom_boxplot()+
  labs(title = "MonthlyCharges distribution")+
      theme_minimal()
ggplot(cus_retention, aes(x=Churn, y=TotalCharges, fill=Churn)) +
  geom_boxplot()+
  labs(title = "TotalCharges distribution")+
      theme_minimal()
```


